Geometric variational problems

### 2020-11-12 NSF2020/Description/specificaim2_ver2.tex

\begin{figure}[h]
  \centering{\includegraphics[width=0.8\textwidth]{Figures/domain.pdf}}
    \caption{\label{fig:flow_map}}
\end{figure}

@@

\subsubsection{Vesicles in background flows}
Numerically solving the Stokes-screened Laplace system~\eqref{SL}
and~\eqref{eq:stokes} is nontrivial, and we will develop efficient,
robust, high-order in time and space algorithms suited to the problem.
The challenges are: {\bf (1)} the equations express a two-way coupling.
This means that the flow changes the shape of the suspended particles,
and in return the geometry-dependent hydrophobic stresses impart a force
on the flow. {\bf (2)} The inputs to the Stokes equations are particle
configurations, forces, and torques. The outputs are the rigid body
translation and angular velocities used to update the particle positions
(see Figure~\ref{fig:flow_map}). Although the underlying equations are
linear, the overall problem is highly nonlinear because the domain is
constantly changing. {\bf (3)} Self-assembly causes the particles to
come into close contact. As a result, an exceptional spatial accuracy is
required to resolve the fields between adjacent particles. Finally, the
physically relevant elastic properties of bilayer become apparent at
large length scales and for large particle-numbers, which increases the
computational complexity of our simulations.

@@

We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established results. \todo{What are the known results?}
Researchers have devised models to enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. Our
preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading  elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.

>>>>

%\begin{figure}[h]
\begin{wrapfigure}[]{r}{0.60\textwidth}
  \centering{\includegraphics[width=0.6\textwidth]{Figures/domain.pdf}}
    \caption{\label{fig:flow_map} \todo[inline]{Missing caption}}
\end{wrapfigure}
%\end{figure}

@@

\subsubsection{Vesicles in background flows}
Numerically solving the Stokes and screened Laplace system~\eqref{SL}
and~\eqref{eq:stokes} is nontrivial, and we will develop efficient,
robust, high-order in time and space algorithms suited to the problem.
The challenges are: {\bf (1)} the equations express a two-way coupling.
This means that the flow changes the shape of the suspended particles,
and in return the geometry-dependent hydrophobic stresses impart a force
on the flow. {\bf (2)} The inputs to the Stokes equations are particle
configurations, forces, and torques. The outputs are the rigid body
translation and angular velocities used to update the particle positions
(see Figure~\ref{fig:flow_map}). Although the underlying equations are
linear, the overall problem is highly nonlinear because the domain is
constantly changing. {\bf (3)} Self-assembly causes the particles to
come into close contact. As a result, an exceptional spatial accuracy is
required to resolve the fields between adjacent particles. Finally, the
physically relevant elastic properties of bilayer become apparent at
large length scales and for large particle-numbers, which increases the
computational complexity of our simulations.

@@

We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. Our
preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.

### 2020-11-12 NSF2020/Description/specificaim3_ver2.tex

Significant challenges to solve BIEs include solving dense linear
systems, developing preconditioning strategies, and developing
quadrature methods for integrands that are nearly singular. These
challenges are described in section~\ref{subsec:NumericalIssues}.
However, numerical errors can still lead to unphysical contact between
bodies. We propose two algorithms to avoid contact in
Section~\ref{subsec:timeStepping}. Then, proposed work for
well-conditioned BIE formulations of fluctuating hydrodynamics are
described in Section~\ref{subsec:fluctuating}.

@@

\begin{wrapfigure}[21]{r}{2.5in}
%\vspace*{+15pt}
\centerline{\includegraphics[width=2.5in]{figures/integrands}}
\vspace*{-13pt}
  \caption{{\footnotesize The Fourier modes of
  $K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ where
  $|\mathbf{x_0}| = 0.99$ and $\mathbf{y} \in \partial\Omega$, and
  $\Omega$ is the unit circle. The trapezoid rule applied to the red
  integrand has large amounts of error, but the trapezoid rule is much
  more accurate when applied to the blue integrand. Since the difference
  between the red and blue curves can be accurately computed with the
  barycentric quadrature rule, the quadrature error applied to layer
  potentials for the screen Laplace equation will be uniformly bounded
  in $\Omega$.}}
\label{fig:integrands}
\end{wrapfigure}
Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. This method requires the
layer potential satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}. The PIs will extend
this quadrature method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature of the trapezoid rule is the sum of the Fourier
amplitudes not captured at the resolution.

>>>>

\begin{wrapfigure}[21]{r}{2.5in}
%\vspace*{+15pt}
\centerline{\includegraphics[width=2.5in]{figures/integrands}}
\vspace*{-13pt}
  \caption{{\footnotesize The Fourier modes of
  $K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ where
  $|\mathbf{x_0}| = 0.99$ and $\mathbf{y} \in \partial\Omega$, and
  $\Omega$ is the unit circle. The trapezoid rule applied to the red
  integrand has large amounts of error, but the trapezoid rule is much
  more accurate when applied to the blue integrand. Since the difference
  between the red and blue curves can be accurately computed with the
  barycentric quadrature rule, the quadrature error applied to layer
  potentials for the screen Laplace equation will be uniformly bounded
  in $\Omega$.}}
\label{fig:integrands}
\end{wrapfigure}
Significant challenges to solve BIEs include solving dense linear
systems, developing preconditioning strategies, and developing
quadrature methods for integrands that are nearly singular. These
challenges are described in section~\ref{subsec:NumericalIssues}.
However, numerical errors can still lead to unphysical contact between
bodies. We propose two algorithms to avoid contact in
Section~\ref{subsec:timeStepping}. Then, proposed work for
well-conditioned BIE formulations of fluctuating hydrodynamics are
described in Section~\ref{subsec:fluctuating}.

@@

Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. This method requires the
layer potential satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}. The PIs will extend
this quadrature method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature of the trapezoid rule is the sum of the Fourier
amplitudes not captured at the resolution.
